PANCHANGE_data_preprocessing

### 2020-08-12 PANCHANGE_data_preprocessing2020-07-15.Rmd
Added chunk to calculate NBR age and to combine NBR cohort and COPING demographic data.

#Packages
Install packages (if they are not available in your version of R)
```{r Installing packages}
#install.packages("summarytools")
#install.packages("tidyverse")
#install.packages("psych")
#install.packages("dplyr")
```

Load packages
```{r Load packages}
library(knitr)
library(summarytools)
library(psych)
library(dplyr)
library(tidyverse)
```

@@

  nbr.dem.raw.id <- nbr.dem.raw %>%
    drop_na(subjectid) %>% # Drop NAs
    distinct(subjectid, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(subjectid, into = c("Sample", "ID.intm"), sep = 6) %>% # Split ID in Sample and Number
    separate(ID.intm, into = "ID", sep = 7) %>%
    mutate(ID = as.numeric(ID),
           startDate.coping = startDate) %>% # Rename startDate
    select(
      Sample, # Sample
      ID, # ID
      startDate.coping, #Date when participant signed up to nbr
#      Age_uncleaned = demographics.how_old_are_you_now.txt, # Age
      Birthyear = demographics.year, # Birthyear
      Gender = demographics.which_gender_do_you_identify_with, # Gender
      Sex = demographics.select_questionnaire_items_medical, #Sex
      Ethnicity = demographics.questions_based_ethnic_origin # Ethnicity
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

@@

nbr.cohort.raw.id <- nbr.cohort.raw %>%
  drop_na(`NBR ID`) %>% # Drop NAs
  distinct(`NBR ID`, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(`NBR ID`, into = c("Sample", "ID.intm"), sep = 6) %>% # Split ID in Sample and Number
  separate(ID.intm, into = "ID", sep = 7) %>%
  mutate(
    ID = as.numeric(ID),
#         startDate.coping = startDate
         ) %>% # Rename startDate
  select(
         Sample, # Sample
         ID, # ID
#         startDate.coping, #Date when participant signed up to NBR COPING
         Age_uncleaned = age, # Age
         Gender = gender, # Gender
#         Sex = demographics.select_questionnaire_items_medical, #Sex
#         EduYrs = demographics.school_education_school_years.txt, # Years in education
#         demographics.college_or_university_degree, # level of education - univeristy
#         demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
#         demographics.o_levelsgcses_or_equivalent, # level of education - gcse
#         demographics.cses_or_equivalent, # level of education - cse
#         demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         Ethnicity = ethnicity # Ethnicity
         ) %>%
  add_numeric(., exclude = exclude_cols_dem) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

```{r nbr copy dem}
if(NBR == TRUE){
dem.raw.id <- nbr.dem.raw.id
}
```

>>>>

#Packages
Install packages (if they are not available in your version of R)
```{r Installing packages}
#install.packages("summarytools")
#install.packages("tidyverse")
#install.packages("psych")
#install.packages("dplyr")
#install.packages("eeptools")
```

Load packages
```{r Load packages}
library(knitr)
library(summarytools)
library(psych)
library(dplyr)
library(tidyverse)
library(eeptools)
```

@@

  nbr.dem.raw.id <- nbr.dem.raw %>%
    drop_na(subjectid) %>% # Drop NAs
    distinct(subjectid, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(subjectid, into = c("Sample", "ID.intm"), sep = 6) %>% # Split ID in Sample and Number
    separate(ID.intm, into = "ID", sep = 7) %>%
    mutate(ID = as.numeric(ID),
           startDate.coping = startDate) %>% # Rename startDate
    select(
      Sample, # Sample
      ID, # ID
      startDate.coping, #Date when participant signed up to nbr
#      Age_uncleaned = demographics.how_old_are_you_now.txt, # Age
      Birthyear_unc = demographics.year, # Birthyear
      Birthmonth = demographics.month, #Birthmonth
      Birthday = demographics.day, #Birthday
      Gender = demographics.which_gender_do_you_identify_with, # Gender
      Sex = demographics.select_questionnaire_items_medical, #Sex
#      Ethnicity = demographics.questions_based_ethnic_origin # Ethnicity
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

```{r nbr dem calculating age from birthday}
if(NBR == TRUE) {
#add 1900 to birthyear
nbr.dem.raw.id$Birthyear_numeric <- nbr.dem.raw.id$Birthyear_unc_numeric + 1900

#create dob as date format
nbr.dem.raw.id$Birthdate <- as.Date(paste(nbr.dem.raw.id$Birthyear_numeric,
                                  nbr.dem.raw.id$Birthmonth_numeric,
                                  nbr.dem.raw.id$Birthday_numeric,
                                  sep = "-"))

head(nbr.dem.raw.id$Birthdate)

#Calculate dem.age(remove NAs as it throws an error)
ages <- age_calc(na.omit(nbr.dem.raw.id$Birthdate), enddate = Sys.Date(), units = "years")

#Create dem.agecolumn
nbr.dem.raw.id$Age_uncleaned<- NA

#Populate with values (excluding NAs)
nbr.dem.raw.id$Age_uncleaned[!is.na(nbr.dem.raw.id$Birthdate)] <- ages

#Round age down
nbr.dem.raw.id$Age_uncleaned <- floor(nbr.dem.raw.id$Age_uncleaned)

#Visualise
head(nbr.dem.raw.id$Age_uncleaned)
head(nbr.dem.raw.id$Birthdate)
}
```

@@

nbr.cohort.raw.id <- nbr.cohort.raw %>%
  drop_na(`NBR ID`) %>% # Drop NAs
  distinct(`NBR ID`, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(`NBR ID`, into = c("Sample", "ID.intm"), sep = 6) %>% # Split ID in Sample and Number
  separate(ID.intm, into = "ID", sep = 7) %>%
  mutate(
    ID = as.numeric(ID),
#         startDate.coping = startDate
         ) %>% # Rename startDate
  select(
         Sample, # Sample
         ID, # ID
#         startDate.coping, #Date when participant signed up to NBR COPING
#         Age_uncleaned = age, # Age
         Gender = gender, # Gender
#         Sex = demographics.select_questionnaire_items_medical, #Sex
#         EduYrs = demographics.school_education_school_years.txt, # Years in education
#         demographics.college_or_university_degree, # level of education - univeristy
#         demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
#         demographics.o_levelsgcses_or_equivalent, # level of education - gcse
#         demographics.cses_or_equivalent, # level of education - cse
#         demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         Ethnicity = ethnicity # Ethnicity
         ) %>%
  add_numeric(., exclude = exclude_cols_dem) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

```{r nbr combine demographics}
if(NBR == TRUE){
dem.raw.id <- merge(nbr.cohort.raw.id, nbr.dem.raw.id, by = "ID")
}
```
